Practical Optional Types for Clojure

### 2015-04-22 formal-model.tex
add reviews

Syntax is given in \figref{main:figure:termsyntax}. Expressions include variables, values,
application, abstractions, conditionals and let expressions.
All binding forms introduce fresh variables.
Values include booleans, \nil{}, class literals, keywords, 
numbers,
constants and closures. 
Value environments map local bindings to values.

@@

As an example, we rewrite
a simple field reference, with the assumptions that
that the class \Point{} has a field $x$ of Java type \Number
and \intaenv {\taenv{}} {p} {\Point{}}.
In rewriting
the expression \fieldexp{x}{p},
{\Point{}} is inferred as the type hint of $p$
and \Number{} is the field type
by \fieldtypeliteral{} (\figref{main:figure:javatyping}).
Now we just plug in the new information into our new expression
$$
\fieldstaticexp{\Point}{\Number}{x}{p}.
$$

@@

The evaluation rules B-Field, B-New and B-Method (\figref{main:figure:javatyping}) simply evaluate their
arguments and call the relevant JVM operation, which do not model---\secref{sec:metatheory}
states our exact assumptions.

@@

To ease presentation, we present \emph{immutable}
multimethods, with syntax and semantics given in \figref{main:figure:mmsyntax}. 
\defmethodliteral{} returns a new extended multimethod
without changing the original multimethod. \egref{example:rep} is now written
\begin{minted}{clojure}
(let [path (defmulti [Any -> (U nil String)] class)]
  (let [path (defmethod path String [x] x)]
    (let [path (defmethod path File [^File x] 
                 (.getPath x))]
      (path "dir/a")))) ;=> "a"
\end{minted}

>>>>

The syntax is given in \figref{main:figure:termsyntax}. Expressions include variables, values,
application, abstractions, conditionals and let expressions.
All binding forms introduce fresh variables.
Values include booleans, \nil{}, class literals, keywords, 
numbers,
constants and closures. 
Value environments map local bindings to values.

@@

As an example, we rewrite
a simple field reference, with the assumptions
that the class \Point{} has a field $x$ of Java type \Number
and \intaenv {\taenv{}} {p} {\Point{}}.
In rewriting
the expression \fieldexp{x}{p},
{\Point{}} is inferred as the type hint of $p$
and \Number{} is the field type
by \fieldtypeliteral{} (\figref{main:figure:javatyping}).
Now we just plug in the new information into our new expression
$$
\fieldstaticexp{\Point}{\Number}{x}{p}.
$$

@@

The evaluation rules B-Field, B-New and B-Method (\figref{main:figure:javatyping}) simply evaluate their
arguments and call the relevant JVM operation, which we do not model---\secref{sec:metatheory}
states our exact assumptions.

@@

To ease presentation, we present \emph{immutable}
multimethods, with syntax and semantics given in \figref{main:figure:mmsyntax}. 
\defmethodliteral{} returns a new extended multimethod
without changing the original multimethod. \egref{example:rep} is now written
\begin{minted}{clojure}
(let [path (defmulti [Any -> (U nil String)] class)]
  (let [path (defmethod path String [x] x)]
    (let [path (defmethod path File [^File x] 
                 (.getPath x))]
      (path (File. "dir/a"))))) ;=> "dir/a"
\end{minted}

### 2015-04-22 overview.tex
add reviews

The typed \clj{fn} supports optional annotations by 
adding
\clj{:-} and a type after a parameter
position
or binding vector 
to annotate parameter types
and return types respectively.
Typed Clojure provides a \clj{check-ns} function to type checks the current namespace.
\clj{number?} is
a Java \java{instanceof} test of \clj{java.lang.Number}.
As in Typed Racket, \clj{U} creates an \emph{untagged union} type, which can take
any number of types.

@@

\clj{inc-map} is like \egref{example:decmap} except the nodes are incremented.
The reasoning is similar, except we only consider one branch (the current method) by
locally considering the current \emph{dispatch value} and reasoning about how it relates
to the \emph{dispatch function}.
For example, 
in the \clj{:do} method we learn the \clj{:op} key is a \clj{:do}, which
narrows our argument type to the \clj{:do} Expr, and similarly for the \clj{:if}
and \clj{:const} methods.

>>>>

The typed \clj{fn} supports optional annotations by 
adding
\clj{:-} and a type after a parameter
position
or binding vector 
to annotate parameter types
and return types respectively.
Typed Clojure provides a \clj{check-ns} function to type check the current namespace.
\clj{number?} is
a Java \java{instanceof} test of \clj{java.lang.Number}.
As in Typed Racket, \clj{U} creates an \emph{untagged union} type, which can take
any number of types.

@@

\clj{inc-leaf} is like \egref{example:decmap} except the nodes are incremented.
The reasoning is similar, except we only consider one branch (the current method) by
locally considering the current \emph{dispatch value} and reasoning about how it relates
to the \emph{dispatch function}.
For example, 
in the \clj{:do} method we learn the \clj{:op} key is a \clj{:do}, which
narrows our argument type to the \clj{:do} Expr, and similarly for the \clj{:if}
and \clj{:const} methods.
